Efficiently stable presentations from error-correcting codes

### 2023-10-10 efficient_stability.tex
Updates from Overleaf

\paragraph{Efficient stability.}
Faced with the apparent difficulty of studying the general question, it is time to refine our focus and formulate the question which we \emph{do} address. To start, let us explicitly note that 
while stability has previously (for the most part) been studied as a question about a group, our formulation makes it a question about a \emph{presentation} of the group. In particular it is known~\cite{gowers2017inverse,de2019operator} that for finite groups $G$ (which are the only groups we consider in this paper) the multiplication table presentation, which has $|G|$ generators for every group element and $|G|^2$ relations for every pairwise product, is $\delta(\eps)=C\eps$-stable\footnote{Here we assume that $\mu_R,\mu_S$ are uniform over the relations and generators of the multiplication table presentation, respectively. Furthermore, the results of~\cite{gowers2017inverse,de2019operator} apply to a notion of \emph{strongly flexible} stability, where the nearby exact representation may act on a space of larger, but not too much larger, dimension.} for some constant $C$ that is independent of the group. This holds even for approximate representations in arbitrary tracial von Neumann algebras, as long as one allows the notion of flexible stability alluded to above.

However the results of~\cite{gowers2017inverse,de2019operator} do not imply the same quantitative stability bounds for $G(h)$ with respect to its defining presentation~\eqref{eq:gh-intro} (even if we assume that $G(h)$ is finite). The reason to prefer the presentation~\eqref{eq:gh-intro} as opposed to the multiplication table presentation of $G(h)$ is that~\eqref{eq:gh-intro} is much more succinct: if $n,k$ are linearly related then it has $O(k)$ generators and relations, as opposed to $2^k$ and $2^{2k}$ respectively. Such a gain is essential when one recalls the interpretation of the presentation $G(h)$ as a ``tester'' for $G$ --- the size of the presentation is then directly related to the amount of \emph{randomness} required by the tester (to sample a random relation); and in computer science applications randomness is seen as an essential resource (we discuss this more below, in the context of quantum computing).

@@

To summarize, approximate stability results for $G(h)$ enable one to obtain structural results about near-optimal strategies in $\game_h$ --- such a result is known as a \emph{rigidity} result in quantum information. A rigidity result about a game called the quantum low-degree test~\cite{natarajan2018low} is at the heart of the proof of $\sf MIP^* = \sf RE$. The analysis of this test requires an efficient stability result for the Weyl-Heisenberg group. Informally, the reason that the stability result needs to be for an efficient presentation of the Weyl-Heisenberg group, as opposed to e.g.\ the multiplication table presentation, is because to obtain the final result it is necessary that the ``complexity'' of the test, or game, is smaller than the ``complexity'' of the object, or group being tested. Of course here we are loose about what we mean by ``complexity,'' and refer to~\cite{ji2021mip,vidickmip} for high-level explanations. The quantum low-degree test is described and analyzed in Section~\ref{sec:pbt}. We end by mentioning an open question: if one was able to obtain $|S_k|,|R_k|=\poly(k)$ while also having $\delta(\eps)=\poly(\eps)$ in the informal theorem stated above, then this result would likely, through the connection we just described, have consequences for the efficient verification of quantum computations in the framework of interactive proof systems---see e.g.~\cite{coladangelo2019verifier,natarajan2023bounding} for a sample of known results in this direction.

>>>>

\paragraph{Efficient stability.}
Faced with the apparent difficulty of studying the general question, it is time to refine our focus and formulate the question which we \emph{do} address. To start, let us explicitly note that 
while stability has previously (for the most part) been studied as a question about a group, our formulation makes it a question about a \emph{presentation} of the group. In particular it is known~\cite{gowers2017inverse,de2019operator} that for finite groups $G$ (which are the only groups we consider in this paper) the multiplication table presentation, which has $|G|$ generators for every group element and $|G|^2$ relations for every pairwise product, is $\delta(\eps)=C\eps$-flexibly stable\footnote{Here we assume that $\mu_R,\mu_S$ are uniform over the relations and generators of the multiplication table presentation, respectively. Furthermore, the results of~\cite{gowers2017inverse,de2019operator} apply to a notion of \emph{flexible} stability, where the nearby exact representation may act on a space of larger, but not too much larger, dimension. See  Definition \ref{def:close}.} for some constant $C$ that is independent of the group. This holds even for approximate representations in arbitrary tracial von Neumann algebras.

\mnote{The following two paragraphs are problematic. The size of $G(h)$ is independent of $n,k$ in general, even if it is finite. The only case in which you can bound the size of $G(h)$ according to $n,k$ is when it is \textbf{abelian}, and we insisted on not assuming that. I feel like we should add all commutation relations before discussing this part.} However the results of~\cite{gowers2017inverse,de2019operator} do not imply the same quantitative stability bounds for $G(h)$ with respect to its defining presentation~\eqref{eq:gh-intro} (even if $G(h)$ is given to be finite). The reason to prefer the presentation~\eqref{eq:gh-intro} as opposed to the multiplication table presentation of $G(h)$ is that~\eqref{eq:gh-intro} is much more succinct: if $n,k$ are linearly related then it has $\textrm{poly}(k)$ generators and relations, as opposed to $2^k$ and $2^{2k}$ respectively. Such a gain is essential when one recalls the interpretation of the presentation $G(h)$ as a ``tester'' for $G$ --- the size of the presentation is then directly related to the amount of \emph{randomness} required by the tester (to sample a random relation); and in computer science applications randomness is seen as an essential resource (we discuss this more below, in the context of quantum computing).

@@

To summarize, approximate stability results for $G(h)$ enable one to obtain structural results about near-optimal strategies in $\game_h$ --- such a result is known as a \emph{rigidity} result in quantum information. A rigidity result about a game called the quantum low-degree test~\cite{natarajan2018low} is at the heart of the proof of $\sf MIP^* = \sf RE$. The analysis of this test requires an efficient stability result for the Weyl-Heisenberg group. Informally, the reason that the stability result needs to be for an efficient presentation of the Weyl-Heisenberg group \mnote{We mix Weyl-Heisenberg with Pauli all the time. In the main text we call them Pauli, and mention that they are a special case of a Heisenberg group. I like the Weyl-Heisenberg name, since we ignore the y operators in this case, and I thought they are formally part of the Pauli group. But, since we already use notations with P throughout, it would be probably easier to change to Pauli here as well. }, as opposed to e.g.\ the multiplication table presentation, is because to obtain the final result it is necessary that the ``complexity'' of the test, or game, is smaller than the ``complexity'' of the object, or group being tested. Of course here we are loose about what we mean by ``complexity,'' and refer to~\cite{ji2021mip,vidickmip} for high-level explanations. The quantum low-degree test is described and analyzed in Section~\ref{sec:pbt}. We end by mentioning an open question: if one was able to obtain $|S_k|,|R_k|=\poly(k)$ while also having $\delta(\eps)=\poly(\eps)$ in the informal theorem stated above, then this result would likely, through the connection we just described, have consequences for the efficient verification of quantum computations in the framework of interactive proof systems---see e.g.~\cite{coladangelo2019verifier,natarajan2023bounding} for a sample of known results in this direction.
